Forage site distributions quantify collective foraging in honey bee colonies

### 2021-11-12 analysis/manuscript/paper.Rmd
latest update to manuscript

@@

Collective decision making provides animals groups with the ability to pool individual information and make informed decisions. Despite the enhanced ability to process information, not every circumstance requires a collective response. In many cases, animal groups do better when individuals prioritise their own personal information over social information provided by conspecifics [@King2007; @Rieucau2011; @Gruter2013]. Identifying the conditions which favour different information use strategies remains a key challenge, both for determining what conditions likely lead to the evolution of collective decision making and for understanding how animal groups might interact in an increasingly anthropomorphic world. Nevertheless, assessing the importance of different information use strategies can be extremely difficult, requiring the actions and decision making of a collective to be compared with that of individual actors. This is fundamentally difficult as making such comparisons demands observations of individual and collective behaviours. A side from limiting the group size that can be studied, in many cases we can observe one, or the other, but not both.

@@

Despite the efficiency benefits the waggle dance reportedly provides, its contribution to colony foraging success remains unclear. In many situations, colonies that are prevented from communicating via dance behaviour achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2004; @Price2019]. In some ecological circumstances, individuals that forgo waiting for dance information in favour of individual search, termed “scouting”, may find food more quickly, and evidence suggests that bees adopt this strategy if the benefits of dance-following drop [@Gruter2008]. Overall, these studies suggest that honey bees use the waggle dance flexibly in response to environmental conditions [@Guter2011; @Gruter2014]. However, it is difficult to empirically probe this relationship further without new methods which can tease apart the extent to which foraging occurs with or without use of the dance.

@@

To what extent social information is used when making foraging decisions, as well as the value of personal information, has been a matter of considerable debate (refs). Much of that debate is informed by simulation studies (ref), and empirical work in which a relatively small number of individuals are observed (ref). Here, we demonstrate that the analysis of waggle dance data can make an important contribution to this debate and provide a methodology to further evaluate how honey bees use this unique behaviour. With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”. Our methodology provides a means of analysing such large data sets and gleaning useful information on the importance of such data to inform the debate about the importance of social information, as well as providing useful colony metrics of foraging activity.

>>>>

Notes from Elli.

The big thing is collective vs individual information use and what caused the evolution of sociality. With honeybees we have an advanced dance language. Wider context question is of how the waggle dance has evolved / what are the conditions that lead to the evolution of the waggle dance, as it has only happened once. This is sort of what Dornhaus, Herman and Visscher are hinting at.

Gruter's stuff shows individuals can vary their use of the dance depending on foraging experience. The hive disorientation stuff shows that a collective can do better without the dance in some environments. Suggests honeybees flexibly use the dance in colony foraging. However, despite this we have no way of quantifying the extent to which honeybees use the dance. We designed a method to quantify its use and fit it against a large empirical data set.

Results are that honeybees in the majority of cases forage collectively but in some sites use personal information. We can then apply the results to evaluate how this varies across environments.

$$
\pi r^2
$$

@@

Collective decision making provides animals groups with the ability to pool individual information and make more informed decisions. Despite an enhanced processing capacity, not every circumstance requires a collective response. In many cases, animal groups do better when individuals prioritise their own personal information over social information provided by conspecifics [@King2007; @Rieucau2011; @Gruter2013]. Identifying the conditions which favour different information use strategies remains a key challenge, both for determining what conditions likely lead to the evolution of collective decision making and for understanding how animal groups might interact in an increasingly anthropomorphic world. Assessing the importance of different information use strategies is not trivial, requiring the actions and decision making of a collective to be compared with that of individual actors. Making such comparisons demands observations of individual and collective behaviours. A side from limiting the group size that can be studied, in many cases we can observe one, or the other, but not both.

@@

Despite the efficiency benefits the waggle dance reportedly provides, its contribution to colony foraging success remains unclear. In many situations, colonies that are prevented from communicating via dance behaviour achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2004; @Price2019]. In some ecological circumstances, individuals that forgo waiting for dance information in favour of individual search, termed “scouting”, may find food more quickly, and evidence suggests that bees adopt this strategy if the benefits of dance-following drop [@Gruter2008]. Whilst direct experiments can show how individuals respond to different scenarios [@Guter2011; @Gruter2014], empirically evaluating a collective foraging strategy in response to different environmental conditions remains a challenge.

@@

To what extent social information is used when making foraging decisions, as well as the value of personal information, has been a matter of considerable debate (refs). Here, we demonstrate that the analysis of waggle dance data can make an important contribution to this debate and provide a methodology to further evaluate how honey bees use this unique behaviour. Recent advances in computational methods have allowed individuals to be tracked within colonies in ways which would be almost impossible to do by hand (Langraff paper). When combined with social network analyses, the use of such technologies have shed new light on many in colony behaviours. By piecing together the behavioural response of individuals into a mathematical description of colony foraging, the model outlined here extends our ability to predict individual behaviour beyond the colony walls and in to the environment. With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”. Our methodology provides a means of analysing such large data sets and gleaning useful information on the importance of such data to inform the debate about the importance of social information, as well as providing useful colony metrics of foraging activity.
